Exploring the Impact of Encouragement and Student Behavior on Academic Effort in Second-year Statistics Courses

### 2024-12-15 paper/paper.qmd
Update "paper.qmd" and "paper.pdf"
change some graphs and some paragraphs

---
title: "Exploring the Impact of Encouragement and Student Behaviour on Academic Effort in Second-year Statistics Courses"
subtitle: "Perception Inspires Effort, Difficulty Diminishes It, and the Influence of Motivational Emails Remains Uncertain"
author: 
  - Shuangyuan Yang
thanks: "Code and data are available at: [https://github.com/Sophiaaa-Y/student-effort-analysis](https://github.com/Sophiaaa-Y/student-effort-analysis.git)."
date: today
date-format: long
abstract: "Student participation is important for academic success. This research employs Bayesian multiple linear regression to examine the effects of task complexity, intrinsic motivation, emotional states, cognitive competence, and types of email interventions on student effort in email-based programs. The findings emphasize the necessity of balancing task difficulty to prevent demotivation, while emotional and motivational factors, especially changes in cognitive competence and interest, substantially increase effort. The findings provide practical advice for developing personalized, economical strategies that combine cognitive and emotional support to enhance academic motivation."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

As the motivation behind students' abilities to overcome challenges, stay motivated [@wang2014staying], and succeed in their academic endeavors, student involvement plays an important role in determining academic achievements [@kahu2018student]. Although its behavioral [@chevalier2014making], cognitive [@mac1991explaining], and emotional [@trautwein2015using] characteristics have all been studied, little is known about how these factors work together within specific interventions. Research on how such factors change in response to particular, customized communication strategies, like email-based interventions, is particularly lacking. To create more specialized and successful strategies to increase student motivation and engagement, it is essential to understand these combined effects. This study examines how email-based academic interventions influence effort through emotional states, cognitive competence, task difficulty, and intrinsic interest.

This study examines the variables affecting how students' study efforts changed in response to an academic intervention sent by email. The intervention's goal was to motivate students to become more interested in statistics and to actively learn about it. The study examines factors like emotional feelings, cognitive competence, perceived difficulty, intrinsic motivation, and email type in order to determine how these determinants all work together to influence differences in effort. The analysis uses more than 600 student records and a Bayesian multiple linear regression model generated by the `rstanarm` package [@rstanarm] in `R` [@citeR] to determine how these factors together affect academic engagement. In contrast to other research that looked at these factors separately [@akbari2020students] [@goes2016incentive], this study provides a thorough framework for understanding how different facets of student engagement and strategies for communication interact to influence behavioral changes under academic background.

@@

This paper is structured as follows. The @sec-intro includes the introduction part of this study, and the @sec-data outlines the data that we used in the study, and describes the sources of the dataset, the behavioral and performance variables selected for analysis, and the reason reasons behind the choice. @sec-model introduces the Bayesian multiple linear regression model that we used to analyze the relationships between students' effort and predictors such as emotional shifts, cognitive competence, difficulty perception, intrinsic interest, and email type, which provide a statistical framework to understand how they influence students' effort. In @sec-results, it shows the findings from the Bayesian multiple linear regression model, including statistical summaries and graphical representations of the model's outcomes, which show details about how different factors contributed positively or negatively to variations in effort. Besides, @sec-discussion interprets these results in the context of email interventions, discusses limitations of the study, and suggests directions for future research to enhance understanding of student engagement. Finally, additional details about the dataset and its processing are included in [Appendix -@sec-data-details], and [Appendix -@sec-model-details] includes diagnostic plots and posterior predictive checks to evaluate the model's reliability. In [Appendix -@sec-pollster-methodology], it reviews survey design, sampling strategies, and weighting strategies. Finally, [Appendix -@sec-idealized-methodology] provides simulations and proposed refinements to enhance the research's context and validate its approach.

@@

While there are additional datasets available that focus on education and learning, they were not used for this study due to constraints. For example, information from Massive Open Online Courses (MOOCs) provide substantial details about student involvement and behavior within MOOCs. It analyzes user activity data like video interactions, quiz completions, assignment submissions, and participation, providing a useful resource for researching online learning behaviors and engagement patterns [@feng2019understanding]. The dataset helps analyze dropout rates, user engagement, and the link between activity levels and learning results. However, because the dataset only covers fully online learning, it is unsuited for researching hybrid learning models such as flipped classrooms, which include in-person and online components. Besides, it lacks attitudinal data, such as changes in student interest, confidence, or perceived worth over time, which are important in evaluating the overall impact of instructional approaches. Furthermore, the dataset's concentration on short-term user engagement contradicts the study's purpose of investigating the long-term impacts of instructional interventions throughout a semester in a blended learning environment. While the MOOC dataset is useful for research into totally online education [@feng2019understanding], it does not fit the study's specific objectives.

@@

The dataset used in this study was sourced from the Nudge Data Repository [@Taback_nudgedata_2022]. It reflects real-world educational phenomena, examining how email interventions influence student attitudes toward statistic. Each observation corresponds to a unique student who received an intervention, with data collected on their behaviors and performance metrics before and after the emails. According to @taback2023randomized, the measurements in this study were carefully designed to align with established behavioral and educational theories, ensuring reliability and validity. Self-reported surveys were used to capture subjective variables, such as changes in emotional states, cognitive competence, and intrinsic motivation. These surveys employed structured Likert-scale items, ranging from "strongly disagree" to "strongly agree", allowing participants to express complex degrees of agreement or disagreement. Emotional states, such as interest and frustration, were assessed before and after the intervention to evaluate the effectiveness of motivational strategies, while cognitive competence was measured by asking participants to rate their confidence in handling statistical tasks. Objective metrics complemented these surveys by providing unbiased data on task complexity and performance outcomes. Task complexity was quantified based on the number of steps required to complete problems and the time spent on each task, while performance outcomes were measured against standardized benchmarks, such as quiz scores and task completion rates. This dual-method approach ensured a thorough understanding of how interventions influenced behavior, capturing both psychological engagement and actual performance. Grounded in theoretical models like self-determination theory and cognitive load theory, the study integrated subjective and objective measures to provide a robust framework for analyzing the predictors of academic effort and engagement.

@@

The data for this study was systematically downloaded, cleaned, analyzed, modeled and visualized using R [@citeR], a extensive statistical programming language. The following packages were used for this study:

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-effort-by-email
#| fig-cap: "Violin Plot Comparing Change in Effort Across Email Types: Interesting Emails Lead to Higher Median Effort Changes with Greater Variability and Outliers"

# Create the violin plot
model_data %>%
  ggplot(aes(x = email, y = `change in effort`, fill = email)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = "#666666", linetype = "solid") +
  geom_boxplot(
    width = 0.15, position = position_dodge(0.9), color = "black",
    alpha = 0.8, outlier.shape = 16, outlier.size = 1.5, outlier.color = "black"
  ) +
  scale_fill_manual(
    values = c("Plain" = "#FF9999", "Interesting" = "#66B3FF"),
    name = "Email Type"
  ) +
  theme_minimal(base_size = 14) +
  labs(
    x = "Email Type",
    y = "Change in Effort"
  ) +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10),
  )
```
@fig-effort-by-email illustrates the distribution of changes in effort among students based on two types of email interventions: "Plain" and "Interesting". Both groups show a median change in effort close to 0, suggesting that the interventions had a minimal overall impact on effort levels for the majority of students. However, the "Interesting" email group demonstrates a broader distribution, indicated by a wider spread in the violin shape, reflecting greater variability in how students responded to this intervention. This group also contains several extreme negative outliers, with effort changes below -3, suggesting that a subset of students experienced a marked decline in effort following the intervention.

@@

Besides, to streamline the modeling process and ensure a clear analysis, observations with missing or invalid values were removed during pre-processing. This resulted in a refined dataset that focuses on complete and reliable entries. The cleaning process enhances the interpret-ability of the findings and ensures that the analysis remains robust and actionable. This careful selection and preparation of variables allow the study to thoroughly investigate the relationships between effort and its potential predictors, providing meaningful information about the efficacy of behavioral nudges in a learning environment.

@@

- $\beta_2$: The coefficient for "change in cognitive competence", reflecting variations in perceived learning ability. This predictor provides valuable information about changes in students' confidence and cognitive self-assessment.

@@

We run the model in R [@citeR] using the `rstanarm` package [@rstanarm]. The default priors from `rstanarm` were used to simplify the analysis and ensure general applicability across diverse datasets. These priors enhance the model's interpret-ability and stability, providing meaningful inferences about the factors influencing changes in student effort while retaining flexibility to adapt to the subtleties of the observed data.

@@

- Intercept, which has mean equals to -1.20 and standard deviation equals to 0.30, provides information on the baseline level of effort. The intrinsic variability in effort that is unrelated to the observed predictors is shown in the negative mean and the credible interval that range from -1.60 to -0.80. This baseline suggests that although interventions are important, there are underlying factors that affect effort in ways that the current model can not account for, such as contextual, situational, or personal characteristics.

@@

- Although the "change in affect" impact is positive, it is not statistically significant because zero is included in its credible interval. This implies that although there is some evidence correlating changes in students' emotional states to more effort, this relationship is not highly reliable. However, the relationship between mental health and academic success is still an important aspect to take into account when designing interventions and conducting additional research.

@@

Emotional states and cognitive abilities interact to influence students' effort levels and determine how they can participate in academic activities in a meaningful and sustained way. Important incentives of motivation include emotional states like increased interest, joy, and decreased frustration; when students experience positive engagement, they are more willing to spend time and effort to their studies, especially in the face of difficulties. More than just increasing motivation, these pleasant feelings produce an internal sense of pleasure, which turns learning into a rewarding activity rather than a painful duty. Cognitive competency also has a similar and equally important role in performance. Students are more persistent and hardworking when they have the self-confidence to take on challenges because they believe in their own abilities and know that their efforts will be successful. This confidence allows individuals to take on challenges bravely, consider them as chances for personal development, and keep going when things become tough. These emotional and cognitive factors have a mutually beneficial relationship, which emphasizes the necessity of thorough approaches that address both aspects at once. An environment that encourages continuous effort can be established by instructors through encouraging curiosity and excitement, lowering stress and frustration, providing clear and helpful feedback, and decomposing difficult work into small steps. These techniques not only improve students' emotional and cognitive readiness but also create a solid foundation for their academic development and long-term success, sustaining the idea that persistence and motivation can be promoted by taking a balanced approach to the head and the heart.

@@

Task difficulty has a complex effect on student engagement, depending on how challenging it is. It can either encourage or discourage participation. Moderately challenging tasks encourage students to participate actively because they balance effort and attainability. By stimulating curiosity, keeping interest, and providing a sense of accomplishment after successful completion, these tasks promote intrinsic motivation and strengthen persistence. But when the level of difficulty is higher than what the students can currently handle, they may failure, causing frustration, feelings of inadequacy, and ultimately disengagement. Excessively challenging tasks have the potential to overwhelm students, weakening their confidence and causing avoidance behaviors. Because of these two effects, it is important to adjust the complexity of work to each individual's needs, making sure that challenges are motivating instead of discouraging. Effective interventions that address this balance include personalized support, which provides targeted feedback and encouragement; instructions, which provides step-by-step guidance to support learning; and adaptive learning systems, which use technology to automatically adjust task difficulty based on student performance in the moment. Effectively adjusting the degree of difficulty allows teachers to create a learning environment that encourages long-term academic persistence, sustains motivation, and promotes student involvement. In addition to supporting students in overcoming challenges, these tactics help them develop the self-confidence and skills that required for sustained academic accomplishment, which ensuring that the advantages of involvement go beyond current task itself.

@@

It is still unclear how customized emails affect student participation, and it seems to have less of an effect than often expected. Usually using inspirational language and acknowledging difficulties, these emails aim to create a feeling of connection and relevance. However, their effectiveness vary greatly. Comments that acknowledge achievements, understand hardships or stress may give students a brief boost in motivation. However, such endeavors might have little or even no impact on the behavior, engagement, or academic achievement of a large number of people. Because the psychological processes that drive engagement are so broad and complicated, like emotional resonance and cognitive effort, it is hard to raise effort through these kinds of interventions. Although tailored emails are accessible and inexpensive, they appear to have a limited impact and diminishing returns, particularly when students get the same messages over and over again, which may cause desensitization or apathy. In addition to increasing significant worries regarding their long-term effectiveness, their contribution to significant academic improvement may have been overestimated. With these unknowns, further investigation is needed to determine which customized emails could actually improve engagement. Using this strategy without such evidence increases the risk of underestimating how hard it is to inspire students and could divert focus from more thorough and efficient ways.

@@

A number of limitations should be considered when analyzing the results of this study. First of all, the dataset lacks important outside factors that are known to have significant effects on student engagement like parental support [@yang2023parental], peer pressure [@wang2018friends], and classroom dynamics [@skinner2012developmental]. An insufficient understanding of the factors influencing shifts in effort might result from their absence. Secondly, the use of self-reported method for emotional states and cognitive competence may lead to response biases such as social desirability and inaccurate self-assessment, which might affect the connections that are observed. Thirdly, the study makes the assumption that predict variables and response variable have linear correlations, which may oversimplify the complex relationships between environmental, cognitive, and emotional factors. In addition, the results are derived from a particular academic and cultural environment, which restricts their generality to other educational environments or demographics. Lastly, because the study only looks at short-term changes in student effort, it doesn't address sustainability over time or the possible long-term effects on academic performance. These limitations emphasize the need for more sophisticated analytical methods, a wider variety of variables, and a study of the persistence of observed changes over time and in different circumstances.

@@

# Histogram with density line for Change in Effort
p1 <- model_data %>%
  ggplot(aes(x = `change in effort`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "red", color = "white", alpha = 0.8) +
  geom_density(color = "black", size = 1) + # Add density curve
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Histogram with density line for Change in Affect
p2 <- model_data %>%
  ggplot(aes(x = `change in affect`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "orange", color = "white", alpha = 0.8) +
  geom_density(color = "black", size = 1) + # Add density curve
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Histogram with density line for Change in Cognitive Competence
p3 <- model_data %>%
  ggplot(aes(x = `change in cognitive competence`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "darkgreen", color = "white", alpha = 0.8) +
  geom_density(color = "black", size = 1) + # Add density curve
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Histogram with density line for Change in Difficulty
p4 <- model_data %>%
  ggplot(aes(x = `change in difficulty`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "purple", color = "white", alpha = 0.8) +
  geom_density(color = "black", size = 1) + # Add density curve
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Histogram with density line for Change in Interest
p5 <- model_data %>%
  ggplot(aes(x = `change in interest`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white", alpha = 0.8) +
  geom_density(color = "black", size = 1) + # Add density curve
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

p1
p2
p3
p4
p5
```

@@

In @fig-ppcheck-and-posteriorvs-prior-1 we implement a posterior predictive check. This shows how well the Bayesian multiple linear model aligns with the observed data. The plot compares the actual observed data *y* with the replicated data *y_rep* generated from the posterior predictive distribution. The lines of posterior predictions closely match the observed data distribution, demonstrating that the model accurately captures the variability and central tendencies of the data. This indicates a good model fit and suggests that the model effectively predicts the outcomes based on the included predictors.

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck-and-posteriorvs-prior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

@@

```{r}
#| label: fig-posterior
#| fig-cap: "The posterior distributions for all model parameters"
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| fig.width: 8

@@

A significant strength of this study lies in its randomized controlled trial design, which is widely regarded as the gold standard for evaluating causal relationships. By randomly assigning participants to intervention and control groups, the study effectively minimized confounding variables, allowing for a clearer assessment of the intervention's direct impact [@etikan2017sampling]. The stratification by cumulative grade point average (CGPA) further ensured that groups were balanced on key characteristics, increasing the reliability of the results. Additionally, the intervention itself-personalized emails—is both cost-effective and scalable, making it an appealing option for institutions seeking to implement large-scale initiatives to enhance student engagement [@timmons2003email].

@@

To accurately measure the intervention's impact, data should be collected at multiple points during the semester, providing a detailed view of how students' attitudes, engagement, and perceptions evolve over time. This approach moves beyond the limitations of relying only on pre-course and post-course surveys. Periodic surveys every 2–3 weeks can capture short-term changes in students' attitudes, interest, and confidence. These surveys could include Likert-scale questions, such as "How confident do you feel about applying the statistics concepts learned this week?" or "How relevant were this week's emails to your learning?" Open-ended questions might ask students to describe what they found helpful or what could be improved. This feedback would help identify trends, such as increased confidence in certain topics or recurring challenges in others.

>>>>

---
title: "Exploring the Impact of Encouragement and Student Behavior on Academic Effort in Second-year Statistics Courses"
subtitle: "Perception Inspires Effort, Difficulty Diminishes It, and the Influence of Motivational Emails Remains Uncertain"
author: 
  - Shuangyuan Yang
thanks: "Code and data are available at: [https://github.com/Sophiaaa-Y/student-effort-analysis](https://github.com/Sophiaaa-Y/student-effort-analysis.git)."
date: today
date-format: long
abstract: "Student participation is important for academic success. This research employs Bayesian multiple linear regression to examine the effects of task complexity, intrinsic motivation, emotional states, cognitive competence, and types of email interventions on student effort in email-based programs. The findings emphasize the necessity of balancing task difficulty to prevent demotivation, while emotional and motivational factors, especially changes in cognitive competence and interest, substantially increase effort. The findings provide practical advice for developing personalized, economical strategies that combine cognitive and emotional support to enhance academic motivation."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

As the motivation behind students' abilities to overcome challenges, stay motivated [@wang2014staying], and succeed in their academic endeavors, student involvement plays an important role in determining academic achievements [@kahu2018student]. Although its behavioral [@chevalier2014making], cognitive [@mac1991explaining], and emotional [@trautwein2015using] characteristics have all been studied, little is known about how these factors work together within specific interventions. Research on how such factors change in response to customized communication strategies, like email-based interventions, is particularly lacking. To create more specialized and successful strategies to increase student motivation and engagement, it is essential to understand these combined effects. This study examines how email-based academic interventions influence effort through emotional states, cognitive competence, task difficulty, and intrinsic interest.

This study examines the variables affecting how students' study efforts changed in response to an academic intervention sent by email. The intervention's goal was to motivate students to become more interested in statistics and to actively learn about it. The study examines factors like feelings, cognitive competence, perceived difficulty, intrinsic motivation, and email type in order to determine how these determinants all work together to influence differences in effort. The analysis uses more than 600 student records and a Bayesian multiple linear regression model generated by the `rstanarm` package [@rstanarm] in `R` [@citeR] to determine how these factors together affect academic engagement. In contrast to other research that looked at these factors separately [@akbari2020students] [@goes2016incentive], this study provides a thorough framework for understanding how different facets of student engagement and strategies for communication interact to influence behavioral changes under academic background.

@@

This paper is structured as follows. The @sec-intro introduces the study, and the @sec-data describes the data, its sources, variables, and reasoning behind the selection. @sec-model details the Bayesian multiple regression model analyzing predictors of effort, such as emotional shifts, cognitive competence, difficulty perception, interest, and email type. The @sec-results presents findings, including statistical summaries and visualizations, while @sec-discussion interprets these results, discusses limitations, and suggests future research. Appendices include additional data details [Appendix -@sec-data-details], diagnostic plots and posterior predictive checks [Appendix -@sec-model-details], survey methodologies [Appendix -@sec-pollster-methodology], and simulations with refinements [Appendix -@sec-idealized-methodology].

@@

While there are additional datasets available that focus on education and learning, they were not used for this study due to constraints. For example, information from Massive Open Online Courses (MOOCs) provides substantial details about student involvement and behavior within MOOCs. It analyzes user activity data like video interactions, quiz completions, assignment submissions, and participation, providing a useful resource for researching online learning behaviors and engagement patterns [@feng2019understanding]. The dataset helps analyze dropout rates, user engagement, and the link between activity levels and learning results. However, because the dataset only covers fully online learning, it is unsuited for researching hybrid learning models such as flipped classrooms, which include in-person and online components.

Besides, it lacks attitudinal data, such as changes in student interest, confidence, or perceived worth over time, which are important in evaluating the overall impact of instructional approaches. Furthermore, the dataset's concentration on short-term user engagement contradicts the study's purpose of investigating the long-term impacts of instructional interventions throughout a semester in a blended learning environment. While the MOOC dataset is useful for research into totally online education [@feng2019understanding], it does not fit the study's specific objectives.

@@

The dataset used in this study was sourced from the Nudge Data Repository [@Taback_nudgedata_2022]. It reflects real-world educational phenomena, examining how email interventions influence student attitudes toward statistic. Each observation corresponds to a unique student who received an intervention, with data collected on their behaviors and performance metrics before and after the emails.

According to @taback2023randomized, the measurements in this study were carefully designed to align with established behavioral and educational theories, ensuring reliability and validity. Self-reported surveys were used to capture subjective variables, such as changes in emotional states, cognitive competence, and intrinsic motivation. These surveys employed structured Likert-scale items, ranging from "strongly disagree" to "strongly agree", allowing participants to express complex degrees of agreement or disagreement. Emotional states, such as interest and frustration, were assessed before and after the intervention to evaluate the effectiveness of motivational strategies, while cognitive competence was measured by asking participants to rate their confidence in handling statistical tasks.

Objective metrics complemented these surveys by providing unbiased data on task complexity and performance outcomes. Task complexity was quantified based on the number of steps required to complete problems and the time spent on each task, while performance outcomes were measured against standardized benchmarks, such as quiz scores and task completion rates. This dual-method approach ensured a thorough understanding of how interventions influenced behavior, capturing both psychological engagement and actual performance. Grounded in theoretical models like self-determination theory and cognitive load theory, the study integrated subjective and objective measures to provide a robust framework for analyzing the predictors of academic effort and engagement.

@@

The data for this study was systematically downloaded, cleaned, analyzed, modeled and visualized using **`R`** [@citeR], an extensive statistical programming language. The following packages were used for this study:

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-effort-by-email
#| fig-cap: "Comparing Change in Effort Across Email Types: Interesting Emails Lead to Higher Median Effort Changes with Greater Variability and Outliers"

# Create the violin plot
model_data %>%
  ggplot(aes(x = email, y = `change in effort`, fill = email)) +
  geom_violin(trim = FALSE, alpha = 0.5, color = "#666666", linetype = "solid") +
  geom_boxplot(
    width = 0.15, position = position_dodge(0.9), color = "black",
    alpha = 0.8, outlier.shape = 16, outlier.size = 1.5, outlier.color = "black"
  ) +
  scale_fill_manual(
    values = c("Plain" = "#FF9999", "Interesting" = "#66B3FF"),
    name = "Email Type"
  ) +
  theme_minimal(base_size = 14) +
  labs(
    x = "Email Type",
    y = "Change in Effort"
  ) +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10),
  )
```
@fig-effort-by-email illustrates the distribution of changes in effort among students based on two types of email interventions: "Plain" and "Interesting". Both groups show a median change in effort close to 0, suggesting that the interventions had a minimal overall impact on effort levels for most students. However, the "Interesting" email group demonstrates a broader distribution, indicated by a wider spread in the violin shape, reflecting greater variability in how students responded to this intervention. This group also contains several extreme negative outliers, with effort changes below -3, suggesting that a subset of students experienced a marked decline in effort following the intervention.

@@

Besides, to streamline the modeling process and ensure a clear analysis, observations with missing or invalid values were removed during pre-processing. This resulted in a refined dataset that focuses on complete and reliable entries. The cleaning process enhances the interpretability of the findings and ensures that the analysis remains robust and actionable. This careful selection and preparation of variables allow the study to thoroughly investigate the relationships between effort and its potential predictors, providing meaningful information about the efficacy of behavioral nudges in a learning environment.

@@

- $\beta_2$: The coefficient for "change in cognitive competence", reflecting variations in perceived learning ability. This predictor provides useful information about changes in students' confidence and cognitive self-assessment.

@@

We run the model in R [@citeR] using the `rstanarm` package [@rstanarm]. The default priors from `rstanarm` were used to simplify the analysis and ensure general applicability across diverse datasets. These priors enhance the model's interpretability and stability, providing meaningful inferences about the factors influencing changes in student effort while retaining flexibility to adapt to the subtleties of the observed data.

@@

- Intercept, which has mean equals to -1.20 and standard deviation equals to 0.30, provides information on the baseline level of effort. The intrinsic variability in effort that is unrelated to the observed predictors is shown in the negative mean and the credible interval that range from -1.60 to -0.80. This baseline suggests that although interventions are important, there are underlying factors that affect effort in ways that the current model cannot account for, such as contextual, situational, or personal characteristics.

@@

- Although the "change in affect" impact is positive, it is not statistically significant because zero is included in its credible interval. This implies that although there is some evidence correlating changes in students' emotional states to more effort, this relationship is not highly reliable. However, the relationship between mental health and academic success is still an important aspect to consider when designing interventions and conducting additional research.

@@

Emotional states and cognitive abilities interact to influence students' effort levels and determine how they can participate in academic activities in a meaningful and sustained way. Important incentives of motivation include emotional states like increased interest, joy, and decreased frustration; when students experience positive engagement, they are more willing to spend time and effort to their studies, especially in the face of difficulties. More than just increasing motivation, these pleasant feelings produce an internal sense of pleasure, which turns learning into a rewarding activity rather than a painful duty.

Cognitive competency also has a similar and equally important role in performance. Students are more persistent and hardworking when they have the self-confidence to take on challenges because they believe in their own abilities and know that their efforts will be successful. This confidence allows individuals to take on challenges bravely, consider them as chances for personal development, and keep going when things become tough.

These emotional and cognitive factors have a mutually beneficial relationship, which emphasizes the necessity of thorough approaches that address both aspects at once. An environment that encourages continuous effort can be established by instructors through encouraging curiosity and excitement, lowering stress and frustration, providing clear and helpful feedback, and decomposing difficult work into small steps. These techniques not only improve students' emotional and cognitive readiness but also create a solid foundation for their academic development and long-term success, sustaining the idea that persistence and motivation can be promoted by taking a balanced approach to the head and the heart.

@@

Task difficulty has a complex effect on student engagement, depending on how challenging it is. It can either encourage or discourage participation. Moderately challenging tasks encourage students to participate actively because they balance effort and attainability. By stimulating curiosity, keeping interest, and providing a sense of accomplishment after successful completion, these tasks promote intrinsic motivation and strengthen persistence. But when the level of difficulty is higher than what the students can currently handle, they may failure, causing frustration, feelings of inadequacy, and ultimately disengagement. Excessively challenging tasks have the potential to overwhelm students, weakening their confidence and causing avoidance behaviors. Because of these two effects, it is important to adjust the complexity of work to everyone's needs, making sure that challenges are motivating instead of discouraging.

Effective interventions that address this balance include personalized support, which provides targeted feedback and encouragement; instructions, which provides step-by-step guidance to support learning; and adaptive learning systems, which use technology to automatically adjust task difficulty based on student performance in the moment. Effectively adjusting the degree of difficulty allows teachers to create a learning environment that encourages long-term academic persistence, sustains motivation, and promotes student involvement. In addition to supporting students in overcoming challenges, these tactics help them develop the self-confidence and skills that required for sustained academic accomplishment, which ensuring that the advantages of involvement go beyond current task itself.

@@

It is still unclear how customized emails affect student participation, and it seems to have less of an effect than often expected. Usually using inspirational language and acknowledging difficulties, these emails aim to create a feeling of connection and relevance. However, their effectiveness varies greatly. Comments that acknowledge achievements, understand hardships or stress may give students a brief boost in motivation. However, such endeavors might have little or even no impact on the behavior, engagement, or academic achievement of a large number of people. Because the psychological processes that drive engagement are so broad and complicated, like emotional resonance and cognitive effort, it is hard to raise effort through these kinds of interventions. Although tailored emails are accessible and inexpensive, they appear to have a limited impact and diminishing returns, particularly when students get the same messages over and over again, which may cause desensitization or apathy. In addition to increasing significant worries regarding their long-term effectiveness, their contribution to significant academic improvement may have been overestimated. With these unknowns, further investigation is needed to determine which customized emails could actually improve engagement. Using this strategy without such evidence increases the risk of underestimating how hard it is to inspire students and could divert focus from more thorough and efficient ways.

@@

A number of limitations should be considered when analyzing the results of this study. First of all, the dataset lacks important outside factors that are known to have significant effects on student engagement like parental support [@yang2023parental], peer pressure [@wang2018friends], and classroom dynamics [@skinner2012developmental]. An insufficient understanding of the factors influencing shifts in effort might result from their absence. Secondly, the use of self-reported method for emotional states and cognitive competence may lead to response biases such as social desirability and inaccurate self-assessment, which might affect the connections that are observed. Thirdly, the study assumes that predict variables and response variable have linear correlations, which may oversimplify the complex relationships between environmental, cognitive, and emotional factors. In addition, the results are derived from a particular academic and cultural environment, which restricts their generality to other educational environments or demographics. Lastly, because the study only looks at short-term changes in student effort, it doesn't address sustainability over time or the possible long-term effects on academic performance. These limitations emphasize the need for more sophisticated analytical methods, a wider variety of variables, and a study of the persistence of observed changes over time and in different circumstances.

@@

# Histogram with density line for Change in Effort
p1 <- model_data %>%
  ggplot(aes(x = `change in effort`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "firebrick", color = "white", alpha = 0.9) +
  geom_density(color = "black", size = 0.7, alpha = 0.6, fill = "pink") +
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Histogram with density line for Change in Affect
p2 <- model_data %>%
  ggplot(aes(x = `change in affect`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "blue", color = "white", alpha = 0.9) +
  geom_density(color = "black", size = 0.7, alpha = 0.6, fill = "pink") +
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Histogram with density line for Change in Cognitive Competence
p3 <- model_data %>%
  ggplot(aes(x = `change in cognitive competence`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "darkgreen", color = "white", alpha = 0.9) +
  geom_density(color = "black", size = 0.7, alpha = 0.6, fill = "pink") +
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Histogram with density line for Change in Difficulty
p4 <- model_data %>%
  ggplot(aes(x = `change in difficulty`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "darkviolet", color = "white", alpha = 0.9) +
  geom_density(color = "black", size = 0.7, alpha = 0.6, fill = "pink") +
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Histogram with density line for Change in Interest
p5 <- model_data %>%
  ggplot(aes(x = `change in interest`, y = ..density..)) +
  geom_histogram(bins = 30, fill = "darkcyan", color = "white", alpha = 0.9) +
  geom_density(color = "black", size = 0.7, alpha = 0.6, fill = "pink") +
  theme_minimal() +
  theme(
    axis.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10)
  ) +
  labs(x = "", y = "Density")

# Combine all plots using patchwork
p1
p2
p3
p4
p5
```

@@

In @fig-ppcheck-and-posteriorvs-prior-1 we implement a posterior predictive check. This shows how well the Bayesian multiple linear model aligns with the observed data. The plot compares the actual observed data $y$ with the replicated data $y_{\text{rep}}$ generated from the posterior predictive distribution. The lines of posterior predictions closely match the observed data distribution, demonstrating that the model accurately captures the variability and central tendencies of the data. This indicates a good model fit and suggests that the model effectively predicts the outcomes based on the included predictors.

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck-and-posteriorvs-prior
#| layout-ncol: 2
#| fig-cap: "Examining How the Model Fits, and Is Affected by the Data"
#| fig-subcap: ["Posterior Prediction Check", "Comparing the Posterior with the Prior"]

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the Convergence of the MCMC Algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

@@

```{r}
#| label: fig-posterior
#| fig-cap: "The Posterior Distributions for All Model Parameters"
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| fig.width: 8

@@

A significant strength of this study lies in its randomized controlled trial design, which is widely regarded as the gold standard for evaluating causal relationships. By randomly assigning participants to intervention and control groups, the study effectively minimized confounding variables, allowing for a clearer assessment of the intervention's direct impact [@etikan2017sampling]. The stratification by cumulative grade point average (CGPA) further ensured that groups were balanced on key characteristics, increasing the reliability of the results. Additionally, the intervention itself, personalized emails, is both cost-effective and scalable, making it an appealing option for institutions seeking to implement large-scale initiatives to enhance student engagement [@timmons2003email].

@@

To accurately measure the intervention's impact, data should be collected at multiple points during the semester, providing a detailed view of how students' attitudes, engagement, and perceptions evolve over time. This approach moves beyond the limitations of relying only on pre-course and post-course surveys. Periodic surveys every 2 to 3 weeks can capture short-term changes in students' attitudes, interest, and confidence. These surveys could include Likert-scale questions, such as "How confident do you feel about applying the statistics concepts learned this week?" or "How relevant were this week's emails to your learning?" Open-ended questions might ask students to describe what they found helpful or what could be improved. This feedback would help identify trends, such as increased confidence in certain topics or recurring challenges in others.
